---
title: "EDS Job Market Report: Fall and Winter"
subtitle: "Data collected between October 2025 and March 2026"
author: "Sofia Sarak"
date: April 11, 2026
format: html
editor: visual
custom-callout:
  green:
    color: "#78a441"
filters:
- custom-callout
---

The following report contains statistics and visualizations summarizing the job openings being posted to online services such as [Indeed](https://www.indeed.com/jobs?q=environmental+data+science&l=&from=searchOnHP&vjk=b9ae5b9d8d4a16b4), [Conservation Job Board](https://www.conservationjobboard.com/), [CalCareers](https://calcareers.ca.gov/CalHRPublic/Search/JobSearchResults.aspx#kw=environmental%20data), and others. Keywords such "environmental data science", "environmental analyst", and "climate analyst" (just to name a few, though a wide range were used) were searched for, and roles pertaining to environmental data analysis, planning, or communication were pulled and re-posted on the [BrenConnect website](https://brenconnect.bren.ucsb.edu/s/group/0F9Kg0000004CHNKA2/meds-2026-jobinternship-opportunities). 

This report summarizes the listings reposted and data collected between October 2025 and March 2026, and offers a brief comparison between the fall (October through early January) and winter (January through March) quarters. **In total, there are 234 job listings summarized in this report, with 104 pertaining to the fall and the other 130 to the winter.** Example job listings from the winter quarter are also included throughout.

### MEDS 2026 Cohort

The Master of Environmental Data Science (MEDS) program at the Bren School teaches data techniques such as statistical modeling, machine learning, and data visualization. Core skills include R, Python, SQL and version control in GitHub, and therefore students typically aim for jobs that utilize these programming languages and practices. Students are also interested in positions that involve GIS/geospatial methods.

Although most, if not all, positions that exemplified the skills taught in the MEDS program were posted, the job search was somewhat customized to the preferences of the current cohort. This was likely most influential when considering the location of the position, as the current class is particularly interested in California, the West Coast of the United States, Texas, and large cities such as New York City and Washington, D.C, with some international and remote interest as well.

These factors guide the job openings posted on BrenConnect, and therefore the results presented in this report.

```{r}
#| echo: false
#| message: false
#| warning: false
# Load necessary packages
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)
library(patchwork)
library(maps)
library(geomtextpath)
```

```{r}
#| echo: false
#| message: false
#| warning: false

# Load fall quarter data
fall <- read_csv(here("data/fall_posts.csv")) %>% 
  clean_names() %>%
  
  # Denote that these data come from the fall quarter
  mutate(season = "fall")


# Load winter quarter data
winter <- read_csv(here("data/winter_posts.csv")) %>% 
  clean_names() %>%
  
  # Fix data types
  mutate(yearly_salary_min = as.numeric(yearly_salary_min),
         yearly_salary_max = as.numeric(yearly_salary_max)) %>% 
  
  # Denote that these data come from the winter quarter
   mutate(season = "winter",
          
          # Fix observations to match standards
          sector = case_when(
            sector == "International Organization" ~ "Federal Government",
            TRUE ~ sector))

# Join into one
all <- bind_rows(fall, winter)

# Clean joined data frame

all_clean <- all %>% 
  
  # Apply geographic groupings for summaries and viz
  mutate(region = case_when(
    city == "Washington, DC" ~ "Washington,D.C.",
    
    state == "AB" ~ NA,
    
    state == "AK" ~ "Alaska",
    
    state == "HI" ~ "Hawaii",
    
    state %in% c("Multiple", "TX/NM") ~ NA,
    
    # West Coast
    state %in% c("CA", "OR", "WA") ~ "West Coast",
    
    # Southwest
    state %in% c("AZ", "NM", "TX", "NV") ~ "Southwest",
    
    # Rocky Mountains
    state %in% c("CO", "UT", "WY", "MT", "ID") ~ "Rocky Mountains",
    
    # Great Plains
    state %in% c("ND", "SD", "NE", "KS", "OK", "IA", "MO") ~ "Great Plains",
    
    # Great Lakes
    state %in% c("MN", "WI", "MI", "IL", "IN", "OH") ~ "Great Lake",
    
    # Mideast
    state %in% c("NY", "NJ", "PA", "DE", "MD", "DC") ~ "Mideast",
    
    # New England
    state %in% c("ME", "NH", "VT", "MA", "RI", "CT") ~ "New England",
    
    # Southeast
    state %in% c("VA", "WV", "KY", "TN", "NC", "SC", "GA",
                 "FL", "AL", "MS", "AR", "LA") ~ "Southeast",
    
    # Default (keeps state abbreviation if unmatched)
    TRUE ~ state
  )) %>% 
  
  # Fix some position types
  mutate(position_type = str_replace(position_type, "reseacher", "researcher"),
         position_type = str_replace(position_type, "specialist", "data specialist"),
         position_type = str_replace(position_type, "data data specialist", "data specialist"),
         position_type = str_replace(position_type, "machine learning data specialist", "machine learning specialist"))

specialist <- fall$position_type[str_detect(unique(fall$position_type), "specialist")]
```

```{r}
#| echo: false
#| message: false
# Create theme for all visualizations
report_theme <- function(){
  theme_classic() +
  theme(axis.title.y = element_blank())}

bren_blue <- "#003660"
bren_green <- "#78a441"
ivory <- "#F5F9E9"
```

# October 2025 through March 2026

## Sector

The Career Development Team classifies alumni employment into seven categories. Following this system, we find that the majority of position listing were in **consulting** (27% of all posts). The other sectors break down as:

-   Corporate: 21%
-   Non-Profit/NGO: 15%
-   State Government: 12%
-   Local Government: 12%
-   Research/Education: 11%
-   Federal Government: 2%

```{r}
#| echo: false
all_clean %>% 
  group_by(sector) %>% 
  summarize(count = n()) %>% 
ggplot(aes(x = reorder(sector, count), y = count)) +
  geom_col(color = bren_blue, fill = alpha(bren_blue, 0.7)) + # position = "dodge"
  geom_text(aes(label = count), hjust = 1.5,
            color = ivory) +
  labs(y = " ") +
  coord_flip() +
  report_theme() +
  #scale_x_discrete(position = "top") +
  #scale_y_reverse() +
  theme(
   axis.line.x = element_blank(),
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank()
  )
```

::: {.green collapse="true" icon="false"}
## Position in Corporate: Grids Analyst \@ BloombergNEF

**Location:** New York, NY

**Description:** BloombergNEF is a strategic research provider covering global commodity markets and the disruptive technologies shaping the future of the energy transition. They are hiring a grids analyst to work on their power grid team, which tracks, analyzes, and forecasts the spending, supply chains, and technology of the US grid. The analyst will focus on the delivery of transmission grid expansion and modernization, taking a data-driven approach.

**Desired Qualifications:**

-   Education: Bachelor's degree in Science, Technology, Engineering or Mathematics.

-   Demonstrated interest in and enthusiasm for the power or energy sector and the role it plays in our society and economy.

-   Experience with Python coding.

-   Preferred:

    -   A Master’s degree in a related field – such as Engineering, Physics or other sciences.

    -   Prior experience working in or analyzing power system fundamentals, transmission and distribution networks, emerging technologies or energy policy.

    -   Experience working on large-scale data projects requiring use of SQL or other database management solutions.

**Details:** Full-time; \$75K - \$80K salary

**Date Posted (to BrenConnect):** March 17, 2026
:::

## Environmental Field

Along with variety in sectors, roles reflected a wide range of environmental niches. The top three pertained to geospatial work (15% of all positions), water/coastal resources (14%), and sustainability (14%). It should be noted that most consulting positions, unless a specific environment was listed, were categorized under resource management.

```{r}
#| echo: false

all_clean %>% 
  
   mutate(environmental_field = case_when(
    is.na(environmental_field) ~ "other data analysis",
    TRUE ~ environmental_field)) %>% 
  
  group_by(environmental_field) %>% 
  summarize(count = n()) %>% 
  
ggplot(aes(x = count, y = reorder(environmental_field, count))) + 
  geom_bar(stat = "identity",
           color = bren_blue,
           fill = alpha(bren_blue, 0.7)) +
  geom_text(aes(label = count), hjust = 1.5,
            color = ivory) +
  labs(x = " ") +
  report_theme() +
   theme(
   axis.line.x = element_blank(),
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank()
  )
```

## Location

As mentioned, the preference of MEDS 2026 cohort had a big influence on which positions were reposted to BrenConnect, and therefore exist in this dataset. Primarily, there is a very large bias for roles located in California.

```{r}
#| echo: false

# bar graph

stats_region <- all_clean %>%  
  filter(region == "West Coast") %>% 
  group_by(state) %>% 
  summarize(count = n())

location_bar <- all_clean %>% 
filter(!is.na(region)) %>% 
group_by(region) %>% 
  summarize(count = n()) %>% 
ggplot(aes(x = count, y = reorder(region, -count))) + 
  geom_bar(stat = "identity",
           color = bren_blue,
           fill = alpha(bren_blue, 0.8)) + 
  geom_text(aes(label = count, hjust = ifelse(count > 4, 1.5, -1),
            color = ifelse(count > 4, ivory, "black"))) +
  
   scale_color_identity(guide = "none") +
  
  labs(title = " ",
  x =  " ") +
  report_theme() +
  theme(
   axis.line.x = element_blank(),
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank()
  )

location_bar
```

```{r}
#| echo: false
# map of United States

state_counts <- all_clean %>%
  filter(!is.na(state)) %>%
  count(state)

# load in lat and long of US states from map library
us_map <- map_data("state")

# add a column that matches state abbreviation to its name
state_counts <- state_counts %>%
  mutate(
    region = tolower(state.name[match(state, state.abb)])
  )

# join and plot
location_map <- us_map %>%
  left_join(state_counts, by = "region") %>%
  ggplot(aes(long, lat, group = group, fill = n)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(1.3) +
  scale_fill_gradient(
    low = alpha(bren_blue, 0.2),
    high = bren_blue,
    na.value = "grey90"
  ) +
  labs(fill = " ",
       title = " ") +
  theme_void() +
  theme(
    axis.text = element_blank(),
    axis.title = element_blank(),
    axis.ticks = element_blank(),
     plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")
  )

location_map
```

```{r}
#| echo: false
# international?
international <- sum(all_clean$country != "USA")

countries <- unique(all_clean$country)
```

There were a total of 11 international positions posted, in India, Sweden, Canada, Singapore, the Netherlands, Australia, Austria, and the UK. There was also one role in Puerto Rico.

## Position Details

```{r}
#| echo: false
#| output: false

# I chose not to include this summary in the report because I do not think the distinction between position titles such as "data analyst" and "data specialist" are distinct enough

 all_clean %>% 
   drop_na(position_type) %>% 
  group_by(position_type) %>% 
  summarize(count = n()) %>% 
ggplot(aes(x = count, y = reorder(position_type, -count))) + 
  geom_bar(stat = "identity",
           color = bren_blue,
           fill = alpha(bren_blue, 0.8)) +
  geom_text(aes(label = count), hjust = 1.5,
            color = ivory) +
  labs(title = "Position Title",
       x = "Number of Positions") +
  theme_classic() +
  theme(axis.title.y = element_blank(),
        plot.margin = unit(c(0.5, 0.5, 0.5, 0.5), "cm")) 
```

Overall, the majority of jobs were permanent, full-time, and on-site positions. Jobs were classified as "limited term" if they listed specific durations, such as "one-year placement".

```{r}
#| echo: false 
commitment <- all_clean %>% 
  group_by(term) %>% 
  summarize(count = n()) %>% 
  filter(term != "part-time") %>% 
ggplot(aes(x = reorder(term, -count), y = count)) + 
  geom_bar(stat = "identity",
           color = bren_blue,
           fill = alpha(bren_blue, 0.7)) +
  geom_text(aes(label = count), vjust = 1.5,
            color = ivory) +
  labs(title = "Commitment Length",
       y = "Number of Positions") +
  theme_classic() +
  scale_x_discrete(labels = c("Permanent", "Limited Term")) +
  theme(axis.title.x = element_blank()) +
  ylim(0, 200) +
   theme(
   axis.line.y = element_blank(),
   axis.text.y = element_blank(),
   axis.ticks.y = element_blank()
  )


type <- all_clean %>% 
  mutate(job_type = case_when(
    job_type == "short-term project" ~ "internship",
    TRUE ~ job_type)) %>% 
  group_by(job_type) %>% 
  summarize(n = n()) %>% 
ggplot(aes(x = reorder(job_type, -n), y = n)) + 
  geom_bar(stat = "identity",
           color = bren_blue,
           fill = alpha(bren_blue, 0.7)) +
  geom_text(aes(label = n, vjust = ifelse(n > 184, 1.5, -1),
            color = ifelse(n > 184, ivory, "black"))) +
  scale_color_identity(guide = "none") +
  labs(title = "Position Type",
       y = "Number of Positions") +
  theme_classic() +
  scale_x_discrete(labels = c("Full-time", "Internship", "Fellowship", "Part-time")) +
  theme(axis.title.x = element_blank()) +
  ylim(0, 200) +
   theme(
   axis.line.y = element_blank(),
   axis.text.y = element_blank(),
   axis.ticks.y = element_blank()
  )


# * note that short term project was lumped into internship!

modality <- all_clean %>% 
  group_by(modality) %>% 
  summarize(n = n()) %>% 
ggplot(aes(x = reorder(modality, -n), y = n)) + 
  geom_bar(stat = "identity",
           color = bren_blue,
           fill = alpha(bren_blue, 0.7)) +
  geom_text(aes(label = n, vjust = ifelse(n > 60, 1.5, -1),
            color = ifelse(n > 60, ivory, "black"))) +
  
  # tells ggplot to use the color strings literally and supresses legend
  scale_color_identity(guide = "none") +
  
  labs(title = "Modality", y = "Number of Positions") +
  theme_classic() +
   scale_x_discrete(labels = c("On-site", "Hybrid", "Remote", "Remote Option", "Hybrid Option")) +
  theme(axis.title.x = element_blank()) +
  ylim(0, 150) +
   theme(
   axis.line.y = element_blank(),
   axis.text.y = element_blank(),
   axis.ticks.y = element_blank()
  )


commitment / type / modality +
  plot_layout(axes = "collect")
```


::: {.green collapse="true" icon="false"}
## Limited Term Position: Research Data Analyst \@ CalRecycle

**Location:** Sacramento County, CA (hybrid)

**Description:** California’s Department of Resources Recycling and Recovery (CalRecycle) brings together the state’s recycling and waste management programs to move the state towards a circular economy that reduces waste and reuses all materials. The Convenience Zone Unit is seeking a research data analyst to use GIS Map Query and Viewer to independently perform a full range of work. In addition, the incumbent will analyze a wide range of program related issues to advise the unit supervisor in the development of program alternatives, policies and procedures.


**Desired Qualifications:**

-   The ability to use GIS Map Quest and Query

-   The ability to read and understand legislations and regulations, and the knowledge of legislative and budget change processes.

-   Experience and knowledge with geographic principles, geographic information systems, spatial studies and demographic studies

-   Meet the minimum requirements of [Research Data Analyst I](https://eservices.calhr.ca.gov/enterprisehrblazorpublic/Public/ClassSpec/ClassSpecDetail/5729).

**Details:** Full-time; \$4,469 - \$6,9220 monthly pay; 12-month duration, with opportunity for extension

**Date Posted (to BrenConnect):** February 25, 2026
:::

The median compensation of all salaried positions (meaning, excluding part-time work and internships) is **80,000 US dollars**. The breakdown by sector can be seen below.

```{r}
#| echo: false
#| warning: false
all_clean %>% 
  mutate(median_salary = (yearly_salary_min + yearly_salary_max)/2) %>% 
  filter(median_salary != 0,
         job_type != "part-time",
         job_type != "internship") %>% 
  
  ggplot(aes(y = median_salary, x = reorder(sector, median_salary))) + 
  geom_boxplot(color = bren_blue,
           fill = alpha(bren_blue, 0.7)) +
  labs(y = "Annual Compensation (US Dollars)",
       x = " ") +
  theme_minimal() +
  theme(
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        axis.title.y = element_text(margin = margin(r = 8))) +
  
  # styling labels for spacing
  
  scale_x_discrete(labels = c("Non-profit/\nNGO", "Research/\nEducation", 
                              "State\nGovernment", "Consulting", "Federal\nGovernment",
                              "Local\nGovernment", "Corporate")) +
  scale_y_continuous(labels = scales::label_number(scale = 1e-3, suffix = "k"))

```

The corporate sector has the highest outliers in the dataset, with salaries ranging all the way up to \$225,000. The non-profit/NGO sector offers the smallest compensation.

It should be mentioned that not all positions in the dataset are salaried position: if they contained hourly or monthly compensation information, annual compensation was approximated (assuming a 37.5-hour work week). Part-time and internship positions are excluded from these figures.

## Skills and Qualifications

GIS, R, and Python are the primary data analysis skills that employers are looking for. Many positions also specify MS Office tools such as Power BI and Excel.

```{r}
#| echo: false
# Create columns that contain TRUE/FALSE values for hard skills mentioned
all_clean <- all_clean %>% 
  mutate(R_req = stringr::str_detect(all_clean$hard_skills, "R"), 
         Python_req = stringr::str_detect(all_clean$hard_skills, "Python"),
         GIS_req = stringr::str_detect(all_clean$hard_skills, "GIS"),
         Excel_req = stringr::str_detect(all_clean$hard_skills, "Excel"),
         SQL_req = stringr::str_detect(all_clean$hard_skills, "SQL"),
         Power_req = stringr::str_detect(all_clean$hard_skills, "Power BI"))

# Create data frame with the number of positions that mentioned each skill
skills <- c("R", "Python", "GIS", "Excel", "SQL", "Power BI")
skills_sum <- c(sum(all_clean$R_req, na.rm = TRUE), 
                sum(all_clean$Python_req, na.rm = TRUE), 
                sum(all_clean$GIS_req, na.rm = TRUE), 
                sum(all_clean$Excel_req, na.rm = TRUE), 
                sum(all_clean$SQL_req, na.rm = TRUE), 
                sum(all_clean$Power_req, na.rm = TRUE))

skills_count <- data.frame(skills, skills_sum)

# Plot data frame
ggplot(skills_count, aes(x = reorder(skills, skills_sum), y = skills_sum)) +
  geom_col(color = bren_blue,
           fill = alpha(bren_blue, 0.8)) +
  geom_text(aes(label = skills_sum), hjust = 1.5,
            color = ivory) +
  coord_flip() +
  labs(y = " ", 
       x = " ") +
  theme_classic() +
  theme(
   axis.line.x = element_blank(),
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank()
  )
```

It was common for positions to not list a specific coding language, but instead use terms such as "data analysis" or "spatial analysis" to convey their qualifications.

::: {.green collapse="true" icon="false"}
## GIS and R Desired: Restoration Data Management & Analytics Coordinator \@ Mote's International Center for Coral Reef Research & Restoration

**Location:** Summerland Key, Florida

**Description:** Located in the center of Lower Keys and only 24 miles from Key West, Mote’s International Center for Coral Reef Research & Restoration is a fully equipped marine science facility dedicated to marine research and education in the Florida Keys. They are currently seeking a restoration data management & analytics coordinator to actively and closely coordinate with the Program Manager in the management, organization, archival, reporting, and analysis of all Coral Reef Restoration Research data. 

**Desired Qualifications:**

-   M.S. degree in the marine or biological sciences form an accredited college or university.
-   Nationally recognized SCUBA certification (Rescue Diver certification required within 3 months of hire; divemaster or higher preferred).
-   Working knowledge of advanced computing systems such as common photogrammetric software (Agisoft Metashape, Viscore), spatial analysis software (GIS).
-   Experience and working familiarity with common data management and statistical analysis software (MatLab, SPSS, SAS, R).

**Details:** Full-time; Compensation information not listed

**Date Posted (to BrenConnect):** January 30, 2026
:::

# Comparing Fall and Winter

There were some fluctuations in the proportion of job listings that represent each sector, between fall and winter. For example, the percent of consulting roles dropped from 33% to 22%. Other trends include jumps in the proportion of corporate, non-profit/NGO, local government, and federal government roles in the winter.

```{r}
#| echo: false
#| message: false
#| warning: false

all_clean %>% 
  group_by(season, sector) %>% 
  summarize(count = n(),
            percent = ifelse(season == "fall", count/104, count/130)) %>% 
ggplot(aes(x = reorder(sector, percent), y = percent, 
           fill = season, 
           color = season)) + 
  geom_col(position = "dodge") +
  geom_text(aes(label = paste0(round(percent * 100), "%"), 
            vjust = ifelse(season == "fall", 1.5, -0.8)),
            hjust = 1.5,
            color = ivory) +

  scale_fill_manual(values = c(alpha(bren_blue, 0.8), alpha(bren_green, 0.8))) +
  scale_color_manual(values = c(bren_blue, bren_green), guide = "none") +
  
  labs(y = " ",
       fill = " ") +
  report_theme() +
  coord_flip() +
   theme(
   axis.line.x = element_blank(),
   axis.text.x = element_blank(),
   axis.ticks.x = element_blank(),
   legend.position = "top", 
   legend.direction = "horizontal"
  ) +
  guides(fill = guide_legend(reverse = TRUE))
```

::: {.green collapse="true" icon="false"}
## Federal Government Position: Wildland Fire Weather Research Fellow \@ USDA Forest Service

**Location:** Seattle, WA

**Description:** The USDA Forest Service's mission is to help sustain forests and grasslands for present and future generations. There is currently a fellowship opportunity available within their Pacific Northwest Research Station. The fellow will address atmospheric influences on lower intensity fires; the area of research is broadly defined, allowing a significant degree of flexibility.

**Desired Qualifications:**

-   Received or is currently pursuing a master's degree in one of the relevant fields.

-   Previous experience with coupled fire-atmosphere models.

**Details:** Full-time; \$70K – \$90K salary; Appointment will initially be up to one year but may be extended

**Date Posted (to BrenConnect):** January 9, 2026
:::

In terms of position type, the **number limited term roles such as -- internships and fellowships -- increased in the winter**. This may be because hiring employers anticipate an influx of interns in the summer and therefore start their hiring process in early January.

```{r}
#| echo: false
#| warning: false

bren_blue_a <- alpha(bren_blue, 0.8)
bren_green_a <- alpha(bren_green, 0.8)

all_clean %>% 
  group_by(season, term) %>% 
  summarize(count = n(),
            percent = ifelse(season == "fall", count/104, count/130)) %>% 
  filter(term != "part-time") %>% 
ggplot(aes(x = reorder(term, -percent), y = percent)) + 
  geom_bar(stat = "identity",
           aes(fill = season, 
           color = season),
           position = "dodge") +
  geom_text(aes(label = paste0(round(percent * 100), "%"),
                hjust = ifelse(season == "fall", 3.5, -2.5)),
            vjust = 1.5,
            color = ivory) +
  labs(title = " ",
       y = " ",
       fill = " ") +
  theme_classic() +
  scale_x_discrete(labels = c("Permanent", "Limited Term")) +
  theme(axis.title.x = element_blank()) +
  scale_fill_manual(values = c(alpha(bren_blue, 0.8), alpha(bren_green, 0.8))) +
  scale_color_manual(values = c(bren_blue, bren_green), guide = "none") +
   theme(
   axis.line.y = element_blank(),
   axis.text.y = element_blank(),
   axis.ticks.y = element_blank()
  )
```

Other variables were not evaluated as they are more dependent on the roles themselves than on seasonality.

## Conclusion

234 environmental data science positions were retrieved from online job boards between October 28, 2025 through March 27, 2026. They covered roles in all sectors and a wide range of environmental focuses, and varied in salary, modality, commitment length, and position type. The corporate and consulting sectors were the highest paying, and the most popular data analysis qualifications were proficiency in GIS and R.

When comparing fall to winter, it becomes apparent that the number of limited term position listings, such as internships, increase in the winter months. Additionally, roles in federal government were only found in the winter (perhaps related to it being the start of the calendar year).

```{r}
#| echo: false 
#| message: false
#| output: false
commitment <- all_clean %>% 
  group_by(season, term) %>% 
  summarize(count = n()) %>% 
  filter(term != "part-time") %>% 
ggplot(aes(x = reorder(term, -count), y = count)) + 
  geom_bar(stat = "identity",
           aes(fill = season, 
           color = season),
           position = "dodge") +
  geom_text(aes(label = count,
                hjust = ifelse(season == "fall", 3.9, -3)),
            vjust = -.7) +
  labs(title = "Commitment Length",
       y = " ") +
  theme_classic() +
  scale_x_discrete(labels = c("Permanent", "Limited Term")) +
  theme(axis.title.x = element_blank()) +
  ylim(0, 200) +
  scale_fill_manual(values = c(alpha(bren_blue, 0.8), alpha(bren_green, 0.8))) +
  scale_color_manual(values = c(bren_blue, bren_green), guide = "none") +
   theme(
   axis.line.y = element_blank(),
   axis.text.y = element_blank(),
   axis.ticks.y = element_blank()
  )

type <- all_clean %>% 
  mutate(job_type = case_when(
    job_type == "short-term project" ~ "internship",
    TRUE ~ job_type)) %>% 
  group_by(season, job_type) %>% 
  summarize(n = n()) %>% 
ggplot(aes(x = reorder(job_type, -n), y = n)) + 
  geom_bar(stat = "identity",
            aes(fill = season, 
           color = season),
           position = "dodge") +
  geom_text(aes(label = n,
                hjust = ifelse(season == "fall", 2.1, -1.1)),
            vjust = -.7) +
  scale_color_identity(guide = "none") +
  labs(title = "Position Type",
       y = " ") +
  theme_classic() +
  scale_x_discrete(labels = c("Full-time", "Internship", "Fellowship", "Part-time")) +
  theme(axis.title.x = element_blank()) +
  ylim(0, 200) +
  scale_fill_manual(values = c(alpha(bren_blue, 0.8), alpha(bren_green, 0.8))) +
  scale_color_manual(values = c(bren_blue, bren_green), guide = "none") +
   theme(
   axis.line.y = element_blank(),
   axis.text.y = element_blank(),
   axis.ticks.y = element_blank()
  )

# * note that short term project was lumped into internship!

modality <- all_clean %>% 
  group_by(season, modality) %>% 
  summarize(n = n()) %>% 
ggplot(aes(x = reorder(modality, -n), y = n)) + 
  geom_bar(stat = "identity",
            aes(fill = season, 
           color = season),
           position = "dodge") +
  geom_text(aes(label = n,
                hjust = ifelse(season == "fall", 1.9, -1)),
            vjust = -.7) +

  labs(title = "Modality", y = " ") +
  theme_classic() +
   scale_x_discrete(labels = c("On-site", "Hybrid", "Remote", "Remote Option", "Hybrid Option")) +
  theme(axis.title.x = element_blank()) +
  ylim(0, 150) +
  scale_fill_manual(values = c(alpha(bren_blue, 0.8), alpha(bren_green, 0.8))) +
  scale_color_manual(values = c(bren_blue, bren_green), guide = "none") +
   theme(
   axis.line.y = element_blank(),
   axis.text.y = element_blank(),
   axis.ticks.y = element_blank()
  )

commitment / type / modality +
  plot_layout(axes = "collect", guides = "collect")
```

```{r}
#| echo: false
#| message: false
#| output: false
all_clean %>% 
  
   mutate(environmental_field = case_when(
    is.na(environmental_field) ~ "other data analysis",
    TRUE ~ environmental_field)) %>% 
  
  group_by(season, environmental_field) %>% 
  summarize(count = n()) %>% 
  
ggplot(aes(x = count, y = reorder(environmental_field, count))) + 
  geom_bar(stat = "identity",
           aes(fill = season, 
           color = season),
           position = "dodge") +
  #geom_text(aes(label = count, color = season), hjust = -.5) +
  labs(x = "Number of Positions") +
  report_theme() +
   scale_fill_manual(values = c(alpha(bren_blue, 0.8), alpha(bren_green, 0.8))) +
  scale_color_manual(values = c(bren_blue, bren_green), guide = "none") 
```
